# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to download a data set of your choice
# and create 2-5 plots that make use of the techniques from Lesson 3.

# You might use a simple histogram, a boxplot split over a categorical variable,
# or a frequency polygon. The choice is yours!

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. any questions you answered, your observations, and summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# ====================================================================================


install.packages("readxl")
rm(list = ls())

```{r}
library(readxl)
df <- read_excel("indicator Motor vehicles per 1000 population 20100824.xlsx", sheet = 1)

dim(df)
summary(df$`2002.0`)


```


# 3d plt of iris dataset
install.packages("plot3D")
library("plot3D")

install.packages("scatterplot3d", dependencies = TRUE)
library(scatterplot3d)

```{r}
data("iris")
x <- sep.l <- iris$Sepal.Length
y <- pet.l <- iris$Petal.Length
z <- sep.w <- iris$Sepal.Width

scatterplot3d(x = x, y = y, z = z, color = as.numeric(x))
scatterplot3d(x = x, y = y, z = z, color = as.numeric(y))
scatterplot3d(x = x, y = y, z = z, color = as.numeric(z))

```

install.packages('gridExtra')

```{r}

summary(df$`Motor vehicles (not 2-wheelers) per 1,000 population`)

summary(df$`2003.0`)
summary(df$`2004.0`)
summary(df$`2005.0`)
summary(df$`2006.0`)
summary(df$`2007.0`)

p1 <- qplot(data = df, x = df$`2002.0`)
p2 <- qplot(data = df, x = df$`2003.0`)
p3 <- qplot(data = df, x = df$`2004.0`)
p4 <- qplot(data = df, x = df$`2005.0`)
p5 <- qplot(data = df, x = df$`2006.0`)
p6 <- qplot(data = df, x = df$`2007.0`)


library(gridExtra)
g <- grid.arrange(p1, p2, p3, p4, p5, p6, ncol=3)
ggsave('graphs.png', g)

x <- df$`Motor vehicles (not 2-wheelers) per 1,000 population`


qplot(data = df, x = df$`2002.0`, color = factor(`Motor vehicles (not 2-wheelers) per 1,000 population`))
ggsave('graphs1.png')

qplot(data = df, x = df$`2003.0`, color = factor(`Motor vehicles (not 2-wheelers) per 1,000 population`))
ggsave('graphs2.png')

qplot(data = df, x = df$`2004.0`, color = factor(`Motor vehicles (not 2-wheelers) per 1,000 population`))
ggsave('graphs3.png')

qplot(data = df, x = df$`2005.0`, color = factor(`Motor vehicles (not 2-wheelers) per 1,000 population`))
ggsave('graphs4.png')

qplot(data = df, x = df$`2006.0`, color = factor(`Motor vehicles (not 2-wheelers) per 1,000 population`))
ggsave('graphs5.png')

qplot(data = df, x = df$`2007.0`, color = factor(`Motor vehicles (not 2-wheelers) per 1,000 population`))
ggsave('graphs6.png')

summary(df)


s1 <- subset(df, df$`Motor vehicles (not 2-wheelers) per 1,000 population`=='United States', select = -`Motor vehicles (not 2-wheelers) per 1,000 population`)
s2 <- subset(df, df$`Motor vehicles (not 2-wheelers) per 1,000 population`=='India', select = -`Motor vehicles (not 2-wheelers) per 1,000 population`)
s3 <- s1 <- subset(df, df$`Motor vehicles (not 2-wheelers) per 1,000 population`=='China', select = -`Motor vehicles (not 2-wheelers) per 1,000 population`)

scatterplot3d(x = s1, y = s2, z = s3, color = as.numeric(s1))

```

#extrcating letters begining with A
```{r}
x <- data.frame(words=c("apple","orange","coconut","apricot"),var=1:4)
x[grep("^a", x$words),]
```

```{r}
a<-df[grep("^a", df$`Motor vehicles (not 2-wheelers) per 1,000 population`),]

```




```{r}

qplot(data = df, x = df$`2007.0`)

qplot(data = df, x = df$`2002.0`) + geom_histogram()+
  geom_freqpoly(aes(df$`2007.0`))
ggsave('graph1.png')
qplot(data = df, x = df$`2002.0`) + geom_histogram()
ggsave('graph2.png')

```

