  Lesson 4
========================================================

***
```{r}
# clear the workspace
rm(list = ls())
#load ggplot2 library
library(ggplot2)
```
### Scatterplots and Perceived Audience Size
# Notes:

```{r}


pf <- read.csv('pseudo_facebook.tsv', sep='\t')
# pf <- read.delim('pseudo_facebook.tsv')

qplot(x = age, y = friend_count, data = pf) 
qplot(age, friend_count, data = pf) 

```

***


### Scatterplots
Notes:

```{r Scatterplots}
ggplot(aes(x = age, y = friend_count), data = pf) +  geom_point()

```

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +  geom_point() + xlim(13,90)

```

***

#### What are some things that you notice right away?
Response:
- people with less than 25 age have more friends 

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) +  geom_point(alpha = 1/20)

```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +  geom_jitter(alpha = 1/20) + xlim(13,90)

```

#### What do you notice in the plot?
Response:
- people with less than 25 age have more friends
- no of friend from 25 age onward go on decreasing 
- there is a peak around 69 age

***

### Coord_trans()
Notes:

```{r Coord_trans()}
  ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) + 
  xlim(13,90) +  
  coord_trans(y = 'sqrt')

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
  ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h=0)) + 
  xlim(13,90) +  
  coord_trans(y = 'sqrt')

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point()  


ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 1/25, position = position_jitter(h=0)) + 
  xlim(13,90) +  
  coord_trans(y = 'sqrt')

```

***

### Overplotting and Domain Knowledge
Notes:
  
***

### Conditional Means
Notes:
install.packages('dplyr')

```{r}
# test

library(dplyr)
df<-data.frame(N=c(10,20,10,20,30), Pl=c("WO","EI","WO","WO","EI"))
group <- group_by(df, N)
result <- as.data.frame(summarise(group, Pl = n_distinct(Pl)))
result

```



```{r Conditional Means}

library('dplyr')
View(pf)
age_groups <- group_by(pf, age)

summary(pf$age)

pf.fc_by_age <- summarise(age_groups, 
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n=n())

pf.fc_by_age <- arrange(pf.fc_by_age, age)




```

Create your plot!

```{r Conditional Means Plot}
# This programming assignment
# will not be graded, but when you
# submit your code, the assignment
# will be marked as correct. By submitting
# your code, we can add to the feedback
# messages and address common mistakes
# in the Instructor Notes.

# You can assess your work by watching
# the solution video.


# Plot mean friend count vs. age using a line graph.
# Be sure you use the correct variable names
# and the correct data frame. You should be working
# with the new data frame created from the dplyr
# functions. The data frame is called 'pf.fc_by_age'.

# Use geom_line() rather than geom_point to create
# the plot. You can look up the documentation for
# geom_line() to see what it does.

# ENTER ALL OF YOUR CODE TO CREATE THE PLOT BELOW THIS LINE.
# ===========================================================
library(ggplot2)
ggplot(aes(x = age, y = friend_count), data = pf) +  geom_line()

pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n=n()) %>%
  arrange(pf.fc_by_age, age)


head(pf.fc_by_age, 20)

ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +  geom_line()

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

ggplot(aes(x = age, y = friend_count), data = pf) +
  xlim(13,90) +  
  geom_point(alpha = 0.05, position = position_jitter(h=0), color='orange') + 
  coord_trans(y = 'sqrt')+
  geom_line(stat = 'summary', fun.y = quantile,
            fun.args = list(probs = .1), linetype =2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile,
            fun.args = list(probs = .5), color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile,
            fun.args = list(probs = .25), linetype =2)+
  geom_line(stat = 'summary', fun.y = quantile,
            fun.args = list(probs = .9), linetype =2, color = 'blue')

```



```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 0.05, position = position_jitter(h=0), color='orange') + 
  coord_cartesian(xlim=c(13,90), ylim = c(0, 1000)) +
  geom_line(stat = 'summary', fun.y = mean)+
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .1), linetype =2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .5), color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .9), linetype =2, color = 'blue')

```

#### What are some of your observations of the plot?
Response:
- friends_count is more below age 30
- there is a peak at 69 age for friend_count

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor(pf$age, pf$friend_count)

cor.test(pf$age, pf$friend_count, method = 'pearson')

with(pf, cor.test(pf$age, pf$friend_count, method = 'pearson'))

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response: -0.027

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))

with(subset(pf, age <= 70), cor.test(age, friend_count, method = 'pearson'))

with(subset(pf, age <= 70), cor.test(age, friend_count, method = 'spearman'))


```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
library(ggplot2)
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +  geom_point() +
  xlim(0, 1000)+ylim(0,1000)

ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

cor.test(pf$www_likes_received, pf$likes_received, method = 'pearson')

with(subset(pf, pf==quantile(pf$www_likes_received, 0.05) && pf==quantile(pf$likes_received, 0.05) ), cor.test(pf$www_likes_received, pf$likes_received, method = 'pearson'))

     

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)

data("Mitchell")

?Mitchell

```

Create your plot!

```{r Temp vs Month}
library(ggplot2)

ggplot(aes(x = Month, y = Temp), data = Mitchell) +  geom_point()

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}


cor(Mitchell$Temp, Mitchell$Month)

ggplot(aes(x = Month, y = Temp), data = Mitchell) +  geom_point() +
scale_x_continuous(breaks=1:12, limits=c(1, 12))


ggplot(aes(x = Month, y = Temp), data = Mitchell) +  geom_point() +
scale_x_continuous(breaks=1:12, limits=c(1, 12))

range(Mitchell$Month)

ggplot(aes(x = Month, y = Temp), data = Mitchell) +  geom_point() +
scale_x_discrete(breaks=seq(0,203, 12))


```

***

### Making Sense of Data
Notes:

x <- seq(0, 4*pi, pi/20)
y <- cos(x)
qplot(x = x, y = y)
dcor.ttest(x, y)


  
```{r Making Sense of Data}
ggplot(aes(x=(Month%%12),y=Temp), data=Mitchell)+
  geom_point()

```

***

### A New Perspective

What do you notice?
Response:
- cor can correlate linear relations 
- dcor can be useful for cyclic relations

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

pf$age_with_months <- pf$age + (1 - pf$dob_month / 12)
pf$age_with_months <- with(pf, age + (1 - dob_month / 12))

ggplot(aes(x=age,y=friend_count_mean), data=pf.fc_by_age)+
  geom_line()


head(pf.fc_by_age, 10)

pf.fc_by_age[17:19,]


```

***

### Age with Months Means
pf <- read.delim('/datasets/ud651/pseudo_facebook.tsv')

```{r Age with Months Means}

pf <- read.delim('/datasets/ud651/pseudo_facebook.tsv')
pf$age_with_months <-pf$age + (1 - pf$dob_month / 12)
suppressMessages(library(dplyr))

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())

pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)



```

Programming Assignment
```{r Programming Assignment}
pf <- read.delim('/datasets/ud651/pseudo_facebook.tsv')
pf$age_with_months <-pf$age + (1 - pf$dob_month / 12)
suppressMessages(library(dplyr))

# ENTER YOUR CODE BELOW THIS LINE
# ========================================================================
age_groups <- group_by(pf, age_with_months)
pf.fc_by_age_months <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())

pf.fc_by_age_months <- arrange(pf.fc_by_age_months, age_with_months)
head(pf.fc_by_age_months)

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
# Create a new line plot showing friend_count_mean versus the new variable,
# age_with_months. Be sure to use the correct data frame (the one you created
# in the last exercise) AND subset the data to investigate users with ages less
# than 71.

# ENTER YOUR CODE BELOW THIS LINE.
# ===============================================================
pf.fc_by_age_months2 <- subset(pf.fc_by_age_months, pf.fc_by_age_months$age_with_months < 71)
ggplot(aes(x=age_with_months,y=friend_count_mean), data=pf.fc_by_age_months2)+
  geom_line()

# or instead of creating new variable say data = subset.. 

ggplot(aes(x=age_with_months,y=friend_count_mean), data=subset(pf.fc_by_age_months, pf.fc_by_age_months$age_with_months < 71))+
  geom_line()

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

p1 <- ggplot(aes(x=age_with_months,y=friend_count_mean), data=pf.fc_by_age_months)+
  geom_line() +
  geom_smooth()

p2 <- ggplot(aes(x=age_with_months,y=friend_count_mean), data=subset(pf.fc_by_age_months, pf.fc_by_age_months$age_with_months < 71))+
  geom_line()+
  geom_smooth()


p3 <- ggplot(aes(x=round(age/5)*5 ,y=friend_count), data=subset(pf, age < 71))+
  geom_line(stat = "summary", fun.y = mean)

install.packages("gridExtra")
library(gridExtra)
grid.arrange(p1,p2,p3, ncol = 1)

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

