Lesson 4
========================================================

***
```{r}
# clear the workspace
rm(list = ls())
#load ggplot2 library
library(ggplot2)
```
### Scatterplots and Perceived Audience Size
# Notes:

```{r}


pf <- read.csv('pseudo_facebook.tsv', sep='\t')
# pf <- read.delim('pseudo_facebook.tsv')

qplot(x = age, y = friend_count, data = pf) 
qplot(age, friend_count, data = pf) 

```

***


### Scatterplots
Notes:

```{r Scatterplots}
ggplot(aes(x = age, y = friend_count), data = pf) +  geom_point()

```

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +  geom_point() + xlim(13,90)

```

***

#### What are some things that you notice right away?
Response:
- people with less than 25 age have more friends 

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) +  geom_point(alpha = 1/20)

```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +  geom_jitter(alpha = 1/20) + xlim(13,90)

```

#### What do you notice in the plot?
Response:
- people with less than 25 age have more friends
- no of friend from 25 age onward go on decreasing 
- there is a peak around 69 age

***

### Coord_trans()
Notes:

```{r Coord_trans()}
  ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) + 
  xlim(13,90) +  
  coord_trans(y = 'sqrt')

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
  ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h=0)) + 
  xlim(13,90) +  
  coord_trans(y = 'sqrt')

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point()  


ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 1/25, position = position_jitter(h=0)) + 
  xlim(13,90) +  
  coord_trans(y = 'sqrt')

```

***

### Overplotting and Domain Knowledge
Notes:
  
***

### Conditional Means
Notes:
install.packages('dplyr')

```{r}
# test

library(dplyr)
df<-data.frame(N=c(10,20,10,20,30), Pl=c("WO","EI","WO","WO","EI"))
group <- group_by(df, N)
result <- as.data.frame(summarise(group, Pl = n_distinct(Pl)))
result

```



```{r Conditional Means}
library('dplyr')
View(pf)
age_groups <- group_by(pf, age)
summarise(age)

pf.fc_by_age <- summarise(age_groups, 
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n=n())

pf.fc_by_age <- arrange(pf.fc_by_age, age)



pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n=n()) %>%
  arrange(pf.fc_by_age, age)


head(pf.fc_by_age, 20)



```

Create your plot!

```{r Conditional Means Plot}
# This programming assignment
# will not be graded, but when you
# submit your code, the assignment
# will be marked as correct. By submitting
# your code, we can add to the feedback
# messages and address common mistakes
# in the Instructor Notes.

# You can assess your work by watching
# the solution video.


# Plot mean friend count vs. age using a line graph.
# Be sure you use the correct variable names
# and the correct data frame. You should be working
# with the new data frame created from the dplyr
# functions. The data frame is called 'pf.fc_by_age'.

# Use geom_line() rather than geom_point to create
# the plot. You can look up the documentation for
# geom_line() to see what it does.

# ENTER ALL OF YOUR CODE TO CREATE THE PLOT BELOW THIS LINE.
# ===========================================================
library(ggplot2)
ggplot(aes(x = age, y = friend_count), data = pf) +  geom_line()

pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n=n()) %>%
  arrange(pf.fc_by_age, age)


head(pf.fc_by_age, 20)

ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +  geom_line()

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

