Lesson 4
========================================================

***
```{r}
# clear the workspace
rm(list = ls())
#load ggplot2 library
library(ggplot2)

data(diamonds)
summary(diamonds)


levels(diamonds$cut)
levels(diamonds$color)
levels(diamonds$clarity)

 ?diamonds
factor(diamonds$color)

```

```{r}
    n=ncol(diamonds)
    count=0
    for(i in 1:n){
       variable_name<-names(diamonds[i])
       factor<-is.factor(diamonds[[i]])
       if(isTRUE(factor) == TRUE){
        count=count+1 
       }
    }
    print(count)
```

# Create a histogram of the price of
# all the diamonds in the diamond data set.

# TYPE YOUR CODE BELOW THE LINE
# =======================================
```{r}
hist(diamonds$price, col="red")
d <- density(diamonds$price) # returns the density data
plot(d)
summary(diamonds$price)
```

```{r}
sum(diamonds$price < 500)
sum(diamonds$price < 250)
sum(diamonds$price >=15000)

```

# Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.

# There won't be a solution video for this
# question so go to the discussions to
# share your thoughts and discover
# what other people find.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Submit your final code when you are ready.

# TYPE YOUR CODE BELOW THE LINE
# ======================================================================
```{r}
max(diamonds$price)
qplot(diamonds$price, xlim=c(0,10000),binwidth = 50, geom="histogram")
ggsave("priceHistogram.png",device = NULL)
```

# Break out the histogram of diamond prices by cut.

# You should have five histograms in separate
# panels on your resulting plot.

# TYPE YOUR CODE BELOW THE LINE
# ======================================================
```{r}
library(ggplot2)
ggplot(data = diamonds,aes(x=price))+
  geom_histogram()+
  facet_wrap(~cut,ncol = 5)

summary(diamonds$price)

summary(diamonds$price)
by(diamonds$price,diamonds$cut, summary)

by(diamonds$price, diamonds$cut, max)
by(diamonds$price, diamonds$cut, min)
by(diamonds$price, diamonds$cut, median)


```

# In the two last exercises, we looked at
# the distribution for diamonds by cut.

# Run the code below in R Studio to generate
# the histogram as a reminder.

# ===============================================================


# ===============================================================

# In the last exercise, we looked at the summary statistics
# for diamond price by cut. If we look at the output table, the
# the median and quartiles are reasonably close to each other.

# diamonds$cut: Fair
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     337    2050    3282    4359    5206   18570 
# ------------------------------------------------------------------------ 
# diamonds$cut: Good
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     327    1145    3050    3929    5028   18790 
# ------------------------------------------------------------------------ 
# diamonds$cut: Very Good
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     336     912    2648    3982    5373   18820 
# ------------------------------------------------------------------------ 
# diamonds$cut: Premium
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     326    1046    3185    4584    6296   18820 
# ------------------------------------------------------------------------ 
# diamonds$cut: Ideal
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     326     878    1810    3458    4678   18810 

# This means the distributions should be somewhat similar,
# but the histograms we created don't show that.

# The 'Fair' and 'Good' diamonds appear to have 
# different distributions compared to the better
# cut diamonds. They seem somewhat uniform
# on the left with long tails on the right.

# Let's look in to this more.

# Look up the documentation for facet_wrap in R Studio.
# Then, scroll back up and add a parameter to facet_wrap so that
# the y-axis in the histograms is not fixed. You want the y-axis to
# be different for each histogram.

# If you want a hint, check out the Instructor Notes.
  <!-- qplot(x = price, data = diamonds) + facet_wrap(~cut, scales="free_y") + geom_freqpoly(aes(color = cut)) -->
  
```{r}
  qplot(x = price, data = diamonds, color = factor(cut)) + facet_wrap(~cut, scales="free_y")

```

# Create a histogram of price per carat
# and facet it by cut. You can make adjustments
# to the code from the previous exercise to get
# started.

# Adjust the bin width and transform the scale
# of the x-axis using log10.

# Submit your final code when you are ready.

# ENTER YOUR CODE BELOW THIS LINE.
# ===========================================================================

ggplot(data = diamonds,aes(x=price/carat))+
  geom_histogram(binwidth = 0.5)+
  facet_wrap(~cut,ncol = 5) + scale_x_log10()

# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.

# There won't be a solution video for this
# exercise so go to the discussion thread for either
# BOXPLOTS BY CLARITY, BOXPLOT BY COLOR, or BOXPLOTS BY CUT
# to share you thoughts and to
# see what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# =================================================================
```{r}

summary(diamonds$price)
library(ggplot2)
library(gridExtra)
p1<-qplot(x=cut, y =price, data = diamonds, geom ='boxplot')
ggsave('priceHistogram_1.png')

p2<-qplot(x=clarity, y =price, data = diamonds, geom ='boxplot')
ggsave('priceHistogram_2.png')

p3<-qplot(x=color, y =price, data = diamonds, geom ='boxplot')
ggsave('priceHistogram_3.png')

g <- grid.arrange(p1, p2, p3, ncol=3)
ggsave('priceHistogram.png', g)

library(ggplot2)
by(diamonds$price, diamonds$cut, summary)
by(diamonds$price, diamonds$clarity, summary)
by(diamonds$price, diamonds$color, summary)

IQR(subset(diamonds)$price)
  
IQR(subset(diamonds, color == 'D')$price)
IQR(subset(diamonds, color == 'J')$price)
```


```{r}

qplot(x=cut, y =price/carat, data = diamonds, geom ='boxplot')

ggsave('priceHistogram_4.png')
# 
# by(diamonds$price,diamonds$cut, summary)
# 
# by(diamonds$price, diamonds$cut, max)
# by(diamonds$price, diamonds$cut, min)
# by(diamonds$price, diamonds$cut, median)

```

```{r}
library(ggplot2)

ggplot(aes(x = diamonds$carat), data = diamonds) +
  geom_histogram()+
  geom_freqpoly(aes(diamonds$carat), binwidth=0.1)


c1 <- subset(diamonds, diamonds$carat=0.1)
names(diamonds)

car_count <- c(NROW(subset(diamonds, diamonds$carat==0.1)),
NROW(subset(diamonds, diamonds$carat==1.6)),
NROW(subset(diamonds, diamonds$carat==0.3)),
NROW(subset(diamonds, diamonds$carat==2.0)),
NROW(subset(diamonds, diamonds$carat==0.8)),
NROW(subset(diamonds, diamonds$carat==3.0)),
NROW(subset(diamonds, diamonds$carat==1.01)),
NROW(subset(diamonds, diamonds$carat==5.0)))

car_count > 2000

ggplot(aes(x = diamonds$carat), data = diamonds) +
  scale_X_continuous(breaks = 0.1)+
  geom_freqpoly(aes(diamonds$carat), binwidth=0.025)+
  xlim(c(0,6))

```


g <- grid.arrange(p1, p2, p3, ncol=3)
ggsave('priceHistogram.png', g)

```{r}
install.packages("readxl")

```

### Scatterplots and Perceived Audience Size
Notes:


***

### Scatterplots
Notes:

```{r Scatterplots}

```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

```

***

### Overplotting
Notes:

```{r Overplotting}

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}

```

Create your plot!

```{r Conditional Means Plot}

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

