Lesson 5
========================================================

```{r}

rm(list=ls())
pf <- read.csv('pseudo_facebook.tsv', sep='\t')

```

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes: 
- to find if older people can perceive their audience size more precisely 

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4 )

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + geom_line(aes(color=gender), stat = 'summary', fun.y = median)


```
# it is observed from this graph that median friend count for females is more than median friend count for males

```{r}
# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

#   age gender mean_friend_count median_friend_count    n
# 1  13 female          247.2953                 150  207
# 2  13   male          184.2342                  61  265
# 3  14 female          329.1938                 245  834
# 4  14   male          157.1204                  88 1201

# See the Instructor Note for two hints.

# DO NOT DELETE THESE NEXT TWO LINES OF CODE
# ==============================================================
pf <- read.delim('pseudo_facebook.tsv')
suppressMessages(library(dplyr))

# ENTER YOUR CODE BELOW THIS LINE.
# ==============================================================
library(dplyr)
#chain functions together
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(as.numeric(friend_count)),
            median_friend_count = median(as.numeric(friend_count)),
            n = n()) %>% 
  ungroup() %>%
  arrange(age)

head((pf.fc_by_age_gender))


# or

library(dplyr)
age_gender_groups <- group_by(pf, age, gender)

#chain functions together
pf.fc_by_age_gender <- summarise(age_gender_groups,
                          mean_friend_count = mean(as.numeric(friend_count)),
                          median_friend_count = median(as.numeric(friend_count)),
                          n = n()) %>% 
                          filter(!is.na(gender)) %>%
                          ungroup()

arrange(pf.fc_by_age_gender, age)

head((pf.fc_by_age_gender))

```


***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}

library(ggplot2)
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)

```

```{r}
# Create a line graph showing the
# median friend count over the ages
# for each gender. Be sure to use
# the data frame you just created,
# pf.fc_by_age_gender.

# See the Instructor Notes for a hint.

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# =================================================


library(ggplot2)
library(ggplot2)

ggplot(aes(x = age, y = median_friend_count),
       data = pf.fc_by_age_gender)+
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)



```

***

### Thinking in Ratios
Notes:
how many times more firends does the average female user have than male user

***

### Wide and Long Format
Notes:
Reshaping of data is required many times 
eg. long dat to wide data

long data eg. | age | gender | mean_friend_count | median_ _friend_count | n |
wide data eg. | age | male | female |

```{r}
install.packages("tidyr")
library(tidyr)

spread(subset(pf.fc_by_age_gender, 
       select = c('gender', 'age', 'median_friend_count')), 
       gender, median_friend_count)


```


***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)
library(dplyr)
library(tidyr)

pf.fc_by_age_gender.wide <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide)


pf.fc_by_age_gender.wide1 <- dcast(pf.fc_by_age_gender,
                                   age ~ gender,
                                   value.var = 'median_friend_count')
                                   
head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line. Look up the documentation
# for geom_hline to do that. Use the parameter
# linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# =================================================
library(dplyr)
library(tidyr)

pf.fc_by_age_gender.wide <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide)

library(ggplot2)
ggplot(aes(x = age, y = ratio),
       data = pf.fc_by_age_gender.wide)+
  geom_line(stat = 'summary', fun.y = median) +
  geom_hline(yintercept=1, linetype=2)


# or
library(ggplot2)
ggplot(aes(x = age, y = female/male),
       data = pf.fc_by_age_gender.wide)+
  geom_line() +
  geom_hline(yintercept=1, alpha = 0.3 , linetype=2)


```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
# Create a variable called year_joined
# in the pf data frame using the variable
# tenure and 2014 as the reference year.

# The variable year joined should contain the year
# that a user joined facebook.

# See the Instructor Notes for three hints if you get
# stuck. Scroll down slowly to see one hint at a time
# if you would like some guidance.

# This programming exercise WILL BE automatically graded.

# DO NOT ALTER THE CODE BELOW THIS LINE
# ========================================================
pf <- read.delim('/datasets/ud651/pseudo_facebook.tsv')

# ENTER YOUR CODE BELOW THIS LINE.
# ========================================================
pf$year_joined <- 2014 - ceiling(pf$tenure/365)
# or
pf$year_joined <- floor(2014 - pf$tenure/365)

head(pf$year_joined)  
```

***

### Cut a Variable
Notes:

```{r}
## generate data for clinical trial example
clinical.trial <-
    data.frame(patient = 1:100,              
               age = rnorm(100, mean = 60, sd = 8),
               year.enroll = sample(paste("19", 85:99, sep = ""),
                 100, replace = TRUE))
summary(clinical.trial)

## basic usage of cut with a numeric variable
c1 <- cut(clinical.trial$age, breaks = 4)
table(c1)    

## year.enroll is a factor, so must convert to numeric first!
c2 <- cut(as.numeric(as.character(clinical.trial$year.enroll)),
            breaks = 3)
table(c2)

## specify break points explicitly using seq function
## look what seq does  
seq(30, 80, by = 10)
## cut the age variable using the seq defined above
c1 <- cut(clinical.trial$age, breaks = seq(30, 80, by = 10))
## table of the resulting factor           
table(c1)



```


```{r Cut a Variable}
summary(pf$year_joined)

table(pf$year_joined)
?cut

# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

# Note that a parenthesis means exclude the year and a
# bracket means include the year.

#    (0, 5) = 1, 2, 3, 4
#    (0, 5] = 1, 2, 3, 4, 5
#    [0, 5) = 0, 1, 2, 3, 4
#    [0, 5] = 0, 1, 2, 3, 4, 5


# Look up the documentation for cut or try the link
# in the Instructor Notes to accomplish this task.

# DO NOT DELETE THE TWO LINES OF CODE BELOW THIS LINE
# ========================================================================
pf <- read.delim('/datasets/ud651/pseudo_facebook.tsv')
pf$year_joined <- floor(2014 - pf$tenure / 365)

# ENTER YOUR CODE BELOW THIS LINE
# ========================================================================
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))

table(pf$year_joined.bucket)

```

***

### Plotting it All Together
Notes:
```{r}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

```{r Plotting it All Together}
# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.

# You should subset the data to exclude the users
# whose year_joined.bucket is NA.

# If you need a hint, see the Instructor Notes.

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# ===================================================

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}

# Write code to do the following:

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

# As a reminder, the parameter linetype can take the values 0-6:

# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

# This assignment is not graded and
# will be marked as correct when you submit.

# The code from the last programming exercise should
# be your starter code!

# ENTER YOUR CODE BELOW THIS LINE
# ==================================================================
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
    geom_line( stat = 'summary', fun.y = mean, linetype = 2)


```

***

### Friending Rate
Notes:

```{r Friending Rate}
pf1 <- subset(pf, (tenure>=1) )

pf1$frate <- pf1$friend_count /pf1$tenure

summary(pf1$frate)


#or
with( subset(pf, tenure >=1), summary(friend_count / tenure))

```

***

### Friendships Initiated
Notes:

What is the median friend rate? 0.2205 

What is the maximum friend rate? 417.0000 

```{r Friendships Initiated}
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.

# You also need to subset the data to only consider user with at least
# one day of tenure.

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# ========================================================================


ggplot(aes(x = tenure, y = friendships_initiated/tenure),
       data = subset(pf, tenure >=1)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

```{r}
# Instead of geom_line(), use geom_smooth() to add a smoother to the plot.
# You can use the defaults for geom_smooth() but do color the line
# by year_joined.bucket

# ALTER THE CODE BELOW THIS LINE
# ==============================================================================

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:
rm(list = ls())

```{r Histograms Revisited}

yo <- read.csv("yogurt.csv")
str(yo)

yo$id <- factor(yo$id)
str(yo)


library(ggplot2)
ggplot(aes(x=price),data=yo)+geom_histogram()
ggplot(aes(x=price),data=yo)+geom_histogram(binwidth = 10)

qplot(data = yo, x=price, fill = I('#F79420'))

length(unique(yo$price))
table(yo$price)

transform(airquality, Ozone = -Ozone)
transform(airquality, new = -Ozone, Temp = (Temp-32)/1.8)

attach(airquality)
transform(Ozone, logOzone = log(Ozone)) # marginally interesting ...
detach(airquality)



```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.

# One way to do this is using the transform
# function. You can look up the function transform
# and run the examples of code at the bottom of the
# documentation to figure out what it does.

# The transform function produces a data frame
# so if you use it then save the result to 'yo'!

# OR you can figure out another way to create the
# variable.

# DO NOT ALTER THE CODE BELOW THIS LINE
# ========================================================
yo <- read.csv('yogurt.csv')

# ENTER YOUR CODE BELOW THIS LINE
# ========================================================
?transform
yo <- transform(yo, all.purchases = strawberry+blueberry+pina.colada+plain+mixed.berry)

```

***

### Prices over Time
Notes:

```{r Prices over Time}
qplot(x=all.purchases, data = yo, binwidth =1, fill = I('#099DD9'))

# Create a scatterplot of price vs time.

# This will be an example of a time series plot.

# Resolve overplotting issues by using
# techniques you learned in Lesson 4.

# What are some things that you notice?

# ENTER YOUR CODE BELOW THIS LINE
# ================================================

ggplot(aes(x=price, y=time), data=yo) + geom_point()

ggplot(aes(x=time, y=price), data = yo) + geom_jitter(alpha=1/4, shape=21, fill = I('#f79420'))

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
# Set the seed for reproducible results
set.seed(4230)
sample.ids <- sample(yo$id ,16)

ggplot(aes(x=time, y=price), data= subset(yo, id %in% sample.ids)) +
         facet_wrap(~ id) +
         geom_line() +
         geom_point(aes(size = all.purchases), pch =1)



set.seed(1000)
sample.ids <- sample(yo$id ,16)

ggplot(aes(x=time, y=price), data= subset(yo, id %in% sample.ids)) +
         facet_wrap(~ id) +
         geom_line() +
         geom_point(aes(size = all.purchases), pch =1)

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***
```{r}
# install.packages("GGally")
library(GGally)

theme_set(theme_minimal(20))

pf <- read.csv('pseudo_facebook.tsv', sep='\t')

# set the seed for reproducible results
set.seed(1836)
pf_subset <- pf[, c(2:15)]
pf_subset = pf[, c('age', 'dob_year', 'dob_month', 'gender', 'tenure')]
pf_subset <- subset(pf, select = -c(userid, year_joined, year_joined_bucket))

names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset),1000),])

pf1 <- pf_subset[sample.int(nrow(pf_subset),1000),]

cov(pf1$friendships_initiated, pf1$friend_count)
cov(pf1$age, pf1$mobile_likes)



```

### Even More Variables
Notes:

***
```{r}

nci <- read.csv('nci.tsv', sep='\t')

#changing the colnames to produce a nicer plot
#colnames(nci) <- c(1:64)

# melt the data to long format
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c('gene', 'case', 'value')
head(nci.long.samp)

# make the heat map
ggplot(aes(y=gene, x=case, fill=value), data=nci.long.samp) + geom_tile() +
  scale_fill_gradient(colors=colorRampPalette(c('blue', 'red'))(100))


       


```


### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

