Lesson 3
========================================================


***

### What to Do First?
Notes:read the data

***

### Pseudo-Facebook User Data
Notes:read .csv file

```{r Pseudo-Facebook User Data}
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
names(pf)

ggplot(aes(x = pf$dob_day),data = pf) +
  geom_histogram(binwidth = 0.5)  +
  scale_x_continuous(breaks = 1:31)

```

***

#### What are some things that you notice about this histogram?
Response:
- Maximum number of users have their DOB on 1st day of the month.
- Minimum number of users have their DOB on 31st day of the month.



***

### Moira's Investigation
Notes:
- FB users consistently underestimate their audience size of post


***

### Estimating Your Audience Size
Notes:
- guessing their audience size for their posts is just 27% of its true size
***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:
- singing video 

#### How many of your friends do you think saw that post?
Response:
- 450

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:
- 60%
***

### Perceived Audience Size
Notes:

***
### Faceting
Notes: 
facet_wrap(formula)
facet_grid(formula)

```{r Faceting}
ggplot(data = pf,aes(x=dob_day))+
geom_bar()+
scale_x_discrete(breaks = 1:31)+
facet_wrap(~dob_month,ncol = 3)
```

#### Let's take another look at our plot. What stands out to you here?
Response:
-it shows the count of users having DOB on a given day for each month
***

### Be Skeptical - Outliers and Anomalies
Notes:
In econometrics and statistics, a top-coded data observation is one for which data points whose values are above an upper bound are censored.
***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:
- bad data about an extrteme case
***

### Friend Count
Notes:
```{r}
library(ggplot2)
library(ggthemes)
ggplot(data = pf,aes(x=(friend_count)))+ geom_bar()+ xlab("friend_count")


ggplot(aes(x = pf$friend_count), data = pf) +
  geom_histogram()


```



 
#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram()
```

#### How is this plot similar to Moira's first plot?
Response:
it has more friend count in 0-125 range, it goes on decreasing and after 3000 3 outliers can be observed
***

### Limiting the Axes
Notes:


```{r Limiting the Axes}
qplot(x = friend_count, data = pf, xlim = c(0,1000)) 

qplot(x = friend_count, data = pf) +
  scale_x_continuous(limits = c(0,1000))

```

### Exploring with Bin Width
Notes:
guess size was quarter of the actual size
***

### Adjusting the Bin Width
Notes:

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))

qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
  
  
### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) + facet_grid(~gender)

# 
# ggplot(aes(x = friend_count), data = pf) +
#   geom_histogram() +
#   scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
#   facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
ggplot(aes(x = friend_count), data = subset(pf, !is.na(gender))) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:women

#### What's the difference between the median friend count for women and men?
Response:median friend count of women > median friend count of men

#### Why would the median be a better measure than the mean?
Response:median helps in avoiding outliers

***

### Tenure
Notes:

```{r Tenure}
ggplot(aes(x = tenure), data = pf) +
   geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')


ggplot(aes(x = tenure/365), data = pf) +
   geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

ggplot(aes(x = tenure/365), data = pf) +
   geom_histogram(binwidth = 1, color = 'black', fill = '#F79420')


# ggplot(aes(x = tenure/365), data = pf) +
#    geom_histogram(binwidth = 0.25, color = 'black', fill = '#F79420')+
#   scale_x_continuous(breaks = seq(1,7,1), lim = c(0,7))

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
ggplot(aes(x = tenure / 365), data = pf) +
  geom_histogram(color = 'black', fill = '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')

```

***

### User Ages
Notes:

```{r User Ages}
ggplot(aes(x = age), data = pf, binwidth = 1) +
  geom_histogram(fill = '#FF0000')+
  scale_x_continuous(breaks = seq(1, 120, 5)) +
  xlab('Age of the user Facebook') +
  ylab('Number of users of given age')

# ggplot(aes(x = age), data = pf) +
#   geom_histogram(binwidth = 1, fill = '#5760AB') +
#   scale_x_continuous(breaks = seq(0, 113, 5))
```

#### What do you notice?
Response:
- maximum number of users are younger
- there are no users below age 13
***

### The Spread of Memes
Notes:
install.packages('gridExtra')
library(gridExtra)
***

### Lada's Money Bag Meme
Notes:
http://www-personal.umich.edu/~ladamic/projects/

***

### Transforming Data
Notes:

install.packages('gridExtra')
library(gridExtra)

```{r}


# define individual plots
summary(pf$friend_count) 

summary(log10(pf$friend_count))

library(ggplot2)
p1 = qplot(x = friend_count , data = pf) + geom_histogram(fill = '#FF0000') +
  scale_x_continuous(limits = c(-250, 1500))
p2 = qplot(x = log10(friend_count)+1 , data = pf) + geom_histogram(fill = '#00FF00')
#   scale_x_continuous(limits = c(0.30, 3.75))

p3 = qplot(x = sqrt(friend_count) , data = pf)+ geom_histogram(fill = '#0000FF')+
  scale_x_continuous(limits = c(-5, 50))
# arrange plots in grid
grid.arrange(p1, p2, p3, ncol=1)


p1 <- ggplot(aes(x= friend_count), data = pf) + geom_histogram(fill = '#FF5588')
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()
grid.arrange(p1, p2, p3, ncol=1)
```

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
ggplot(aes(x = friend_count), data = pf) + geom_histogram(fill = '#990088')+ scale_x_log10()

```

***


### Frequency Polygons

```{r Frequency Polygons}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Percentage of users with that friend count')

ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10()
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
by(pf$www_likes,pf$gender, sum)


```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(x=gender, y =friend_count, data = subset(pf, !is.na(gender)), geom ='boxplot')

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x=gender, y =friend_count, data = subset(pf, !is.na(gender)), geom ='boxplot', ylim = c(0,1000))

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(x=gender, y =friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom ='boxplot') + 
      coord_cartesian(ylim = c(0,250))
by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:females
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
qplot(x=gender, y =friendships_initiated, 
      data = subset(pf, !is.na(gender)), 
      geom ='boxplot') + 
      coord_cartesian(ylim = c(0,150))
by(pf$friendships_initiated, pf$gender, summary)

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes)

summary(pf$mobile_likes > 0)

mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

sum(pf$mobile_check_in==1)/length(pf$mobile_check_in)*100

```

Response: 64.59

***

### Analyzing One Variable
Reflection:
learnt about various tools for visualization of the data like: qplot, ggplot, scale, facet, boxplot, summary the facebook user data

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!